---
title: "Using nbapredict"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using nbapredict}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `nbapredict` package provides tools for analyzing, visualizing, and predicting NBA player statistics. It combines data cleaning, descriptive statistics, visualization, linear regression, random forest machine learning, model comparison, and hypothesis testing.

This vignette illustrates a typical workflow for using the package.

# Load the package and data

```{r}
library(nbapredict)

data("nba_stats")

head(nba_stats)
```

The included `nba_stats` dataset contains historical NBA player statistics, including season, player name, position, age, team, games played, minutes played, points, rebounds, assists, and points per game.

# Descriptive statistics

The `nba_descriptive_stats()` function summarizes player performance by position.

```{r}
desc <- nba_descriptive_stats(nba_stats)
desc
```

This allows users to compare average points, rebounds, and assists across player positions.

# Visualization

The `nba_plot_analysis()` function creates plots for exploring player statistics.

```{r}
nba_plot_analysis(nba_stats, plot_type = "boxplot")
```

The default boxplot shows the distribution of points per game across player positions.

# Preparing data for modeling

Before building models, the data can be prepared using `nba_prepare_data()`.

```{r}
prepared <- nba_prepare_data(nba_stats)

str(prepared)
```

For demonstration purposes, we use a smaller sample.

```{r}
set.seed(123)
demo_data <- prepared[sample(nrow(prepared), 1000), ]
```

# Linear regression model

The `nba_build_regression_model()` function builds a linear regression model.

```{r}
stat_model <- nba_build_regression_model(demo_data, target = "points")
summary(stat_model)
```

This model predicts total points using player characteristics such as age, minutes played, position, and season.

# Random forest model

The `nba_build_ml_model()` function builds a random forest model.

```{r}
ml_model <- nba_build_ml_model(demo_data, target = "points", ntree = 10)
ml_model
```

The random forest model provides a machine learning approach to predicting player performance.

# Model comparison

The `nba_compare_models()` function compares the statistical model and the machine learning model.

```{r}
comparison <- nba_compare_models(stat_model, ml_model, demo_data)
comparison
```

The output includes RMSE, MAE, and R-squared values for both models.

# Hypothesis testing

The `nba_hypothesis_test()` function performs an ANOVA test to compare player performance across positions.

```{r}
test_result <- nba_hypothesis_test(demo_data, target = "points")
test_result
```

# Advantages and learned points

This package demonstrates several important skills learned in the course:

1. Building a structured R package with standard package files.
2. Writing reusable functions for data analysis.
3. Using roxygen2 documentation to generate help files.
4. Including an example dataset inside an R package.
5. Using `dplyr` for data cleaning and grouped summaries.
6. Using the pipe operator `%>%` to make code more readable.
7. Creating visualizations with `ggplot2`.
8. Writing automated tests with `testthat`.
9. Building statistical models using linear regression.
10. Building machine learning models using random forests.
11. Comparing model performance using RMSE, MAE, and R-squared.
12. Running `devtools::check()` to verify package quality.

Overall, `nbapredict` provides a complete workflow for NBA player statistics analysis, from data preparation to visualization, modeling, comparison, and hypothesis testing.
